A tool set for random number generation on GPUs in R

### 2022-01-12 pkg/gpuRandom/inst/documents/paper1_2021_5/paper1.tex
item sep
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/diseasemapping@3115 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

Below we generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}. The best performance we have seen using \fct{rnormGpu} is above over 100 times faster with $512 \times 128$ work-items than using the standard (and single-threaded) \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size increases.
\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(512 * 128)
R> system.time(rnormGpu(c(10000, 10000), streams = streams, Nglobal = c(512,
+    128), type = "double"))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.045   0.007   0.052
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 256*64)
R> month_gpu <- vclMatrix(month, type = "integer")
R> system.time(result_month <- 
+  clrng::fisher.sim(month_gpu, 1e6, 
+     streams=streams, type="double", 
+     returnStatistics=TRUE,  Nglobal = c(256,64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.317   0.000   0.317
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
R> system.time(resultWeek <- clrng::fisher.sim(week_GPU, 10000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   1.962   0.002   1.966
\end{CodeOutput}
\begin{CodeInput}
R> unlist(resultWeek[c("threshold", "simNum", "counts")])
\end{CodeInput}
\begin{CodeOutput}
## threshold    simNum    counts 
##    -54990  10010624      1350
\end{CodeOutput}
\begin{CodeInput}
R> resultWeek$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.000135
\end{CodeOutput} 
\end{CodeChunk}

@@

\caption{\label{tab:summarycompare}Summary of comparions of Fisher's test simulation on different devices. Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII. Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100.\label{tab:summary}, run-time varies a little bit for each execution.}
\centering
\begin{tabular}[t]{l|l|l|l|l|l}
\hline
\multicolumn{1}{c|}{ } & \multicolumn{2}{c|}{Computer 1} & \multicolumn{2}{c|}{Computer 2} & \multicolumn{1}{c}{ } \\
\cline{2-3} \cline{4-5}
B & Intel 2.5ghz & AMD Radeon & Intel 3.7ghz & NVIDIA V100 & Data\\
\hline
\multicolumn{6}{l}{\textbf{P-value}}\\
\hline
\hspace{1em}1M & 0.403804 & 0.403507 & 0.4035606 & 0.403507 & month\\
\hline
\hspace{1em}10M & 0.0001251 & 0.0001274 & 0.0001202 & 0.0001274 & week\\
\hline
\multicolumn{6}{l}{\textbf{Run-time}}\\
\hline
\hspace{1em}1M & 49.3 & 2.2 & 14.8 & 0.3 & month\\
\hline
\hspace{1em}10M & 327.3 & 10.2 & 91.5 & 1.9 & week\\
\hline
\end{tabular}
\end{table}

>>>>

Below we generate a large-size matrix of 100 million double-precision Gaussian random numbers, and compare the run-time between using \fct{stats::rnorm} and \fct{rnormGpu}. The best performance we have seen using \fct{rnormGpu} is above over 100 times faster with $512 \times 128$ work-items than using the standard (and single-threaded) \fct{stats::rnorm}. The difference in elapsed time becomes larger when matrix size increases.
\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(512 * 128)
R> system.time(rnormGpu(c(10000, 10000), streams = streams, Nglobal = c(512,
+    128), type = "double"))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.060   0.011   0.071
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> streams <- createStreamsGpu(n = 256*64)
R> month_gpu <- vclMatrix(month, type = "integer")
R> system.time(result_month <- 
+  clrng::fisher.sim(month_gpu, 1e6, 
+     streams=streams, type="double", 
+     returnStatistics=TRUE,  Nglobal = c(256,64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   0.318   0.000   0.318
\end{CodeOutput} 
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> week_GPU <- gpuR::vclMatrix(week, type = "integer")
R> system.time(resultWeek <- clrng::fisher.sim(week_GPU, 10000000,
+    streams = streams, type = "double", returnStatistics = TRUE,
+    Nglobal = c(256, 64)))
\end{CodeInput}
\begin{CodeOutput}
##    user  system elapsed 
##   1.925   0.027   1.951
\end{CodeOutput}
\begin{CodeInput}
R> unlist(resultWeek[c("threshold", "simNum", "counts")])
\end{CodeInput}
\begin{CodeOutput}
## threshold    simNum    counts 
##    -54990  10010624      1350
\end{CodeOutput}
\begin{CodeInput}
R> resultWeek$p.value
\end{CodeInput}
\begin{CodeOutput}
## [1] 0.000135
\end{CodeOutput} 
\end{CodeChunk}

@@

\caption{\label{tab:summarycompare}Summary of comparions of Fisher's test simulation on different devices. Computer 1 is equipped with CPU Intel Xenon W-2145 3.7Ghz and AMD Radeon VII. Computer 2 is equipped with VCPU Intel Xenon Skylake 2.5Ghz and VGPU Nvidia Tesla V100 \label{tab:summary}. Run-time varies a little bit for each execution.}
\centering
\begin{tabular}[t]{l|l|l|l|l|l}
\hline
\multicolumn{1}{c|}{ } & \multicolumn{2}{c|}{Computer 1} & \multicolumn{2}{c|}{Computer 2} & \multicolumn{1}{c}{ } \\
\cline{2-3} \cline{4-5}
B & Intel 2.5ghz & AMD Radeon & Intel 3.7ghz & NVIDIA V100 & Data\\
\hline
\multicolumn{6}{l}{\textbf{P-value}}\\
\hline
\hspace{1em}1M & 0.403804 & 0.403507 & 0.4035606 & 0.403507 & month\\
\hline
\hspace{1em}10M & 0.0001251 & 0.0001274 & 0.0001202 & 0.0001274 & week\\
\hline
\multicolumn{6}{l}{\textbf{Run-time}}\\
\hline
\hspace{1em}1M & 49.3 & 2.2 & 14.8 & 0.3 & month\\
\hline
\hspace{1em}10M & 327.3 & 10.2 & 91.5 & 1.9 & week\\
\hline
\end{tabular}
\end{table}
